---
title: "Poisson Population"
subtitle: "Discrete counts with lambda = 1"
format: html
---

```{r}
source("../R/helpers.R")
student_id <- 6599L  # Replace with the last four digits of your student ID.
B <- B_DEFAULT
study <- study_population("poisson", seed = student_id, B = B)
d <- study$diagnostics
```

## Population and prediction

The observations are independent Poisson(lambda = 1) counts. The population is
discrete and right-skewed, with mean and variance equal to 1. I expect the
sample-mean distribution to become smoother and less skewed with \(n\), but the
small rate means that small samples will retain visible discreteness.

## Experimental design and decision rule

I generated `r paste0("B = ", B)` sample means at candidate sizes `r paste(N_VALUES, collapse = ", ")`. The common screen requires absolute skewness ≤ 0.20, absolute
excess kurtosis ≤ 0.30, Q-Q correlation ≥ 0.995, and standardized Q-Q RMSE ≤
0.08, with three consecutive passing values required.

## Results

```{r}
display <- d[, c("n", "simulated_mean", "simulated_sd", "theoretical_sd",
                 "skewness", "excess_kurtosis", "qq_correlation", "qq_rmse",
                 "normal_enough", "stable_pass")]
display$normal_enough <- ifelse(display$normal_enough, "yes", "no")
display$stable_pass <- ifelse(display$stable_pass, "yes", "no")
knitr::kable(display, digits = 3)
```

## Visual evidence

```{r}
plot_n <- if (is.na(study$selected_n)) max(N_VALUES) else study$selected_n
x <- study$samples[[as.character(plot_n)]]
old_par <- par(mfrow = c(1, 2))
hist(x, breaks = 40, probability = TRUE, col = "orchid", border = "white",
     main = paste("Sampling means: n =", plot_n), xlab = "Sample mean")
curve(dnorm(x, mean = study$spec$mean, sd = study$spec$sd_mean(plot_n)),
      add = TRUE, col = "firebrick", lwd = 2)
qqnorm(x, main = paste("Normal Q-Q plot: n =", plot_n), pch = 16, cex = 0.3)
qqline(x, col = "firebrick", lwd = 2)
par(old_par)
```

## Conclusion

The first stable pass is **`r ifelse(is.na(study$selected_n), "no stable value in the investigated range", paste0("n = ", study$selected_n))`**. The result shows how a discrete, moderately skewed population can approach a bell shape through averaging, while its original integer-valued support is not itself normal.

## Reproducibility check

```{r}
reproducibility_check("poisson", n = plot_n, seed_a = student_id,
                      seed_b = 2026L, B = B)
```
