---
title: "Cauchy Population"
subtitle: "A deliberate CLT limitation"
format: html
---

```{r}
source("../R/helpers.R")
student_id <- 6599L  # Replace with the last four digits of your student ID.
B <- B_DEFAULT
study <- study_population("cauchy", seed = student_id, B = B)
d <- study$diagnostics
```

## Population and prediction

The observations are independent Cauchy(location = 0, scale = 1) values. The
Cauchy distribution has no finite population mean or variance, so the usual
finite-variance Central Limit Theorem does not apply. I predict that ordinary
sample means will remain unstable and will not meet the normality criteria.

## Experimental design and decision rule

I generated `r paste0("B = ", B)` sample means at candidate sizes `r paste(N_VALUES, collapse = ", ")`. The common shape screen requires absolute skewness ≤ 0.20,
absolute excess kurtosis ≤ 0.30, Q-Q correlation ≥ 0.995, and standardized Q-Q
RMSE ≤ 0.08, with three consecutive passing values required. For Cauchy data,
the simulated mean and standard deviation are reported descriptively only; they
are not estimates of finite population parameters.

## Results

```{r}
display <- d[, c("n", "simulated_mean", "simulated_sd", "skewness",
                 "excess_kurtosis", "qq_correlation", "qq_rmse",
                 "normal_enough", "stable_pass")]
display$normal_enough <- ifelse(display$normal_enough, "yes", "no")
display$stable_pass <- ifelse(display$stable_pass, "yes", "no")
knitr::kable(display, digits = 3)
```

## Visual evidence

```{r}
plot_n <- max(N_VALUES)
x <- study$samples[[as.character(plot_n)]]
old_par <- par(mfrow = c(1, 2))
hist(x, breaks = 50, probability = TRUE, col = "gray60", border = "white",
     main = paste("Cauchy sample means: n =", plot_n), xlab = "Sample mean")
qqnorm(x, main = paste("Normal Q-Q plot: n =", plot_n), pch = 16, cex = 0.3)
qqline(x, col = "firebrick", lwd = 2)
par(old_par)
```

## Conclusion

The first stable pass is **`r ifelse(is.na(study$selected_n), "no stable value in the investigated range", paste0("n = ", study$selected_n))`**. Within the investigated range, no stable \(n\) met all criteria. This is not evidence that a larger finite \(n\) must eventually work; it is evidence that ordinary sample means are not governed by the usual finite-mean, finite-variance CLT here. The heavy tails also make exact summaries highly seed-sensitive.

## Reproducibility check

```{r}
reproducibility_check("cauchy", n = plot_n, seed_a = student_id,
                      seed_b = 2026L, B = B)
```
