---
title: "Non-identical Bernoulli Process"
subtitle: "Bounded observations with changing probabilities"
format: html
---

```{r}
source("../R/helpers.R")
student_id <- 6599L  # Replace with the last four digits of your student ID.
B <- B_DEFAULT
study <- study_population("nonidentical_bernoulli", seed = student_id, B = B)
d <- study$diagnostics
```

## Population and prediction

For a sample of size \(n\), the observations are independent Bernoulli values,
but their probabilities change linearly from 0.50 to 0.90. They are therefore
not identically distributed. The average success probability is 0.70. Because
the observations remain bounded, I expect a bell shape eventually, although the
usual iid explanation of the CLT does not apply directly.

## Experimental design and decision rule

I generated `r paste0("B = ", B)` sample means at candidate sizes `r paste(N_VALUES, collapse = ", ")`. I use the same four criteria for approximate normality:
absolute skewness ≤ 0.20, absolute excess kurtosis ≤ 0.30, Q-Q correlation ≥
0.995, and standardized Q-Q RMSE ≤ 0.08. Three consecutive passes define a
stable selection. The theoretical standard deviation uses the sum of the
individual Bernoulli variances rather than an iid formula.

## Results

```{r}
display <- d[, c("n", "simulated_mean", "simulated_sd", "theoretical_sd",
                 "skewness", "excess_kurtosis", "qq_correlation", "qq_rmse",
                 "normal_enough", "stable_pass")]
display$normal_enough <- ifelse(display$normal_enough, "yes", "no")
display$stable_pass <- ifelse(display$stable_pass, "yes", "no")
knitr::kable(display, digits = 3)
```

## Visual evidence

```{r}
plot_n <- if (is.na(study$selected_n)) max(N_VALUES) else study$selected_n
x <- study$samples[[as.character(plot_n)]]
old_par <- par(mfrow = c(1, 2))
hist(x, breaks = 40, probability = TRUE, col = "cadetblue", border = "white",
     main = paste("Non-identical means: n =", plot_n), xlab = "Sample mean")
curve(dnorm(x, mean = study$spec$mean, sd = study$spec$sd_mean(plot_n)),
      add = TRUE, col = "firebrick", lwd = 2)
qqnorm(x, main = paste("Normal Q-Q plot: n =", plot_n), pch = 16, cex = 0.3)
qqline(x, col = "firebrick", lwd = 2)
par(old_par)
```

## Conclusion

The first stable pass is **`r ifelse(is.na(study$selected_n), "no stable value in the investigated range", paste0("n = ", study$selected_n))`**. This case separates identical distribution from independence: changing probabilities alter the variance and centering calculation, but the bounded independent contributions can still produce an approximately normal average.

## Reproducibility check

```{r}
reproducibility_check("nonidentical_bernoulli", n = plot_n, seed_a = student_id,
                      seed_b = 2026L, B = B)
```
