---
title: "Custom Age-at-Death Distribution"
subtitle: "A discrete two-component population built from the supplied reference code"
format: html
---

```{r}
source("../R/helpers.R")
student_id <- 6599L  # Replace with the last four digits of your student ID.
B <- B_DEFAULT
study <- study_population("mixture", seed = student_id, B = B)
d <- study$diagnostics
full_grid_path <- "../results/tables/full_grid_summary.csv"
full_grid_n <- NA_integer_
if (file.exists(full_grid_path)) {
  full_grid_summary <- read.csv(full_grid_path, stringsAsFactors = FALSE)
  full_grid_n <- full_grid_summary$smallest_n_judged_normal[
    full_grid_summary$population == study$spec$label
  ][1]
}
```

## Population and prediction

This analysis uses the custom age-at-death distribution in
`Inference/custom_age_at_death_distribution_monte_carlo.html`. The reference
code defines integer ages from 0 through 100, adds an early-life exponential
weight and an adult bell-shaped weight centered at age 65, and normalizes the
combined weights so they sum to one. The project implements that same generator
through `custom_age_at_death_distribution()` and `rdeath()` in `R/helpers.R`.

This is a synthetic teaching distribution designed to reproduce the supplied
shape; it is not a fitted historical mortality table. Because it has finite
mean and variance, the CLT predicts eventual convergence of sample means, but
the two-component shape should make small-sample normality less immediate than
for symmetric populations.

## Reference population shape

```{r}
death_distribution <- custom_age_at_death_distribution()
death_mean <- with(death_distribution, sum(age * probability))
death_sd <- with(
  death_distribution,
  sqrt(sum((age - death_mean)^2 * probability))
)
probability_under_20 <- with(
  death_distribution,
  sum(probability[age < 20])
)
probability_60_or_older <- with(
  death_distribution,
  sum(probability[age >= 60])
)
theoretical_median <- death_distribution$age[
  which(cumsum(death_distribution$probability) >= 0.5)[1]
]

plot(
  death_distribution$age,
  death_distribution$percent,
  type = "l",
  lwd = 3,
  col = "darkorange",
  xlab = "Age at death",
  ylab = "Percent of deaths",
  main = "Supplied custom age-at-death distribution"
)
grid(col = "gray90", lty = 2)
lines(death_distribution$age, death_distribution$percent,
      col = "darkorange", lwd = 3)
```

The distribution has theoretical mean **`r sprintf("%.2f", death_mean)` years**
and standard deviation **`r sprintf("%.2f", death_sd)` years**. It assigns
approximately **`r sprintf("%.2f", 100 * probability_under_20)`%** of its
probability to ages below 20 and **`r sprintf("%.2f", 100 * probability_60_or_older)`%**
to ages 60 and older. These values are calculated from the normalized discrete
probabilities, not estimated from the Monte Carlo sample means.

## Large-population simulation check

The reference HTML also simulates a population of 100,000 people before
studying repeated sample means. I reproduce that check here with the same
`rdeath()` generator, using the project’s recorded student seed (`r student_id`)
instead of the reference lesson’s demonstration seed, so the simulated
summaries remain consistent with the rest of this project.

```{r}
set.seed(student_id)
number_of_people <- 100000L
simulated_ages <- rdeath(number_of_people, death_distribution)

simulation_summary <- data.frame(
  statistic = c(
    "Mean age at death",
    "Median age at death",
    "Standard deviation",
    "Percent below age 20",
    "Percent age 60 or older"
  ),
  theoretical = c(
    death_mean,
    theoretical_median,
    death_sd,
    100 * probability_under_20,
    100 * probability_60_or_older
  ),
  simulated = c(
    mean(simulated_ages),
    median(simulated_ages),
    sd(simulated_ages),
    100 * mean(simulated_ages < 20),
    100 * mean(simulated_ages >= 60)
  )
)
simulation_summary$difference <-
  simulation_summary$simulated - simulation_summary$theoretical
knitr::kable(simulation_summary, digits = 3)
```

The simulated mean, spread, and age-group percentages should be close to their
theoretical counterparts, while the median is reported as a descriptive check
of the population’s two-component shape. Small differences are expected from
finite simulation error.

## Experimental design and decision rule

I generated `r paste0("B = ", B)` sample means at candidate sizes `r paste(N_VALUES, collapse = ", ")`. The common criteria are absolute skewness ≤ 0.20, absolute
excess kurtosis ≤ 0.30, Q-Q correlation ≥ 0.995, and standardized Q-Q RMSE ≤
0.08. Three consecutive passing values are required.

## Results

```{r}
display <- d[, c("n", "simulated_mean", "simulated_sd", "theoretical_sd",
                 "skewness", "excess_kurtosis", "qq_correlation", "qq_rmse",
                 "normal_enough", "stable_pass")]
display$normal_enough <- ifelse(display$normal_enough, "yes", "no")
display$stable_pass <- ifelse(display$stable_pass, "yes", "no")
knitr::kable(display, digits = 3)
```

## Visual evidence

```{r}
plot_n <- if (is.na(study$selected_n)) max(N_VALUES) else study$selected_n
x <- study$samples[[as.character(plot_n)]]
old_par <- par(mfrow = c(1, 2))
hist(x, breaks = 40, probability = TRUE, col = "darkseagreen", border = "white",
     main = paste("Age-at-death means: n =", plot_n), xlab = "Sample mean")
curve(dnorm(x, mean = study$spec$mean, sd = study$spec$sd_mean(plot_n)),
      add = TRUE, col = "firebrick", lwd = 2)
qqnorm(x, main = paste("Normal Q-Q plot: n =", plot_n), pch = 16, cex = 0.3)
qqline(x, col = "firebrick", lwd = 2)
par(old_par)
```

## Conclusion

The first stable pass is **`r ifelse(is.na(study$selected_n), "no stable value in the investigated range", paste0("n = ", study$selected_n))`** on the project’s candidate grid. The separate exhaustive scan of every integer from 2 through 500 reports **`r ifelse(is.na(full_grid_n), "no stable value in the investigated range", paste0("n = ", full_grid_n))`**. These values answer slightly different questions: the candidate-grid result is the first stable pass among the published summary sizes, while the exhaustive result refines the boundary to consecutive integers.

The early-life decline and adult peak create a non-normal population shape that averaging must smooth. The result is evidence for this synthetic death distribution under the stated criteria—not a universal threshold for real mortality data.

## Reproducibility check

```{r}
reproducibility_check("mixture", n = plot_n, seed_a = student_id,
                      seed_b = 2026L, B = B)
```
